Case for Support

### 2021-11-16 application/Je-S_sections.tex
Updates from Overleaf

Scientists studying brain and behaviour dedicate the majority of their time to the design of experimental software protocols that control hardware and to data analysis, with least amount spent on data acquisition itself. Moreover, hundreds of neuroscience research groups worldwide develop their own software tools for experimental control and data analysis in different programming languages, leading to inefficiencies resulting from time-loss spent on writing software (often by individuals lacking programming expertise), the re-invention of software with the same functionality and concomitant lack of experimental reproducibility across laboratories.

Here we propose to provide a community software tool that will dramatically increase the efficiency of experimental control and data analysis by providing a new set of functionalities to an existing software platform, Bonsai. Bonsai is an open-source reactive visual programming environment that naturally orchestrates complex networks of sensors and hardware into workflows and allows integration of multiple programming languages into its framework. Thus far, Bonsai has been adopted by hundreds of scientists worldwide to provide interactive experimental control in behavioural and brain sciences. In this proposal, we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural datasets, and to create an open-access platform for software sharing. Bonsai’s enhanced functionality will enable new types of research via online experimental control, speed up discovery and improve efficiency by (i) providing access to such tools into laboratories lacking expertise, (ii) reducing the need to reinvent the same tools in multiple labs and (iii) standardising the data processing streams, thus increasing reproducibility across laboratories. We believe this effort will enable and accelerate new discoveries in how the brain generates behaviour.

@@

To understand the brain, scientists aim to explain how animal behaviour relates to the activity in neural circuits. This necessitates design and precise control of behavioural experiments wherein experimental animals perform particular tasks while experimenters either record or manipulate neural activity in candidate neural circuits. Such experiments require data acquisition software that integrates and controls hardware from multiple recording devices (cameras, electrodes, sensors), and analysis tools that can interpret large and complex behavioural and neural datasets. Progress is held back by the lack of standardised tools for rapid design and implementation of experimental protocols, and the difficulty of integrating state-of-the-art data processing and neuroinformatics into custom experimental designs. The fields of behavioural and brain sciences have consequently suffered from both inefficiency and poor reproducibility, due to disparate data acquisition and analysis solutions created independently across laboratories.   
To address these challenges, we propose  to  extend,  enhance,  maintain  and  support \textbf{Bonsai},  a  fully  integrated  software  environment  to enable cutting-edge reproducible laboratory systems neuroscience experiments in animal models, with a particular emphasis on machine-intelligence-enabled, real-time neuroinformatics methods. While Bonsai is already adopted by hundreds of scientists worldwide to provide interactive experimental control, we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural data (video-based analysis of behavioural motifs, real-time and off-line analysis of neural signals), and create an open-access platform for software sharing and integration with multiple programming languages. Enhancing Bonsai's ecosystem will enable new types of research, increase user base, and dramatically improve efficiency and reproducibility of research across behavioural and brain sciences.
[44 chars over]

>>>>

Scientists studying brain and behaviour dedicate the majority of their time designing experiments and analysing the data, with least amount spent on data acquisition itself, which may impact the quality of data. Moreover, hundreds of neuroscience research groups worldwide develop their own experimental and analytical tools, most using different programming languages, leading to  inefficiencies in data sharing and analysis, and impacting reproducibility (i.e. how easy it is for someone else to do the same experiment). 
Here we propose to provide the neuroscience community a software tool that will dramatically increase the efficiency of experimental control and data analysis. We will do so by developing a new set of functionalities to an existing software platform, Bonsai. Bonsai is an open-source visual programming environment that naturally orchestrates complex networks of sensors and hardware into workflows and allows integration of multiple programming languages into its framework. Thus far, Bonsai has been adopted by hundreds of scientists worldwide to provide interactive experimental control in behavioural and brain sciences. In this proposal, we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural datasets, and to create an open-access platform for software sharing. Bonsai’s enhanced functionality will enable new types of research via online experimental control, speed up discovery and improve efficiency by (i) providing access to such tools into laboratories lacking expertise, (ii) reducing the need to reinvent the same tools in multiple labs and (iii) standardising the data processing streams, thus increasing reproducibility across laboratories. We believe this effort will enable and accelerate new discoveries in how the brain generates behaviour.

@@

To understand the brain, scientists aim to explain how animal behaviour relates to the activity in neural circuits. This necessitates design and precise control of behavioural experiments wherein experimental animals perform particular tasks while experimenters either record or manipulate neural activity in candidate neural circuits. Such experiments require data acquisition software that integrates and controls hardware from multiple recording devices (cameras, electrodes, sensors), and analysis tools that can interpret large and complex behavioural and neural datasets. Progress is held back by the lack of standardised tools for rapid design and implementation of experimental protocols, and the difficulty of integrating state-of-the-art data processing and neuroinformatics into custom experimental designs. The fields of behavioural and brain sciences have consequently suffered from both inefficiency and poor reproducibility, due to disparate data acquisition and analysis solutions created independently across laboratories.   
To address these challenges, we propose  to  extend,  enhance,  maintain  and  support \textbf{Bonsai},  a  fully  integrated  software  environment  to enable cutting-edge reproducible laboratory systems neuroscience experiments in animal models, with a particular emphasis on machine-intelligence-enabled, real-time neuroinformatics methods. While Bonsai is already adopted by hundreds of scientists worldwide to provide interactive experimental control, we aim to extend Bonsai’s functionality with a toolbox of online and offline Machine Intelligence tools for analysis of behavioural and neural data (video-based analysis of behavioural motifs, real-time and off-line analysis of neural signals), and create an open-access platform for software sharing and integration with multiple programming languages. Enhancing Bonsai's ecosystem will enable new types of research, increase and diversify user base, and dramatically improve efficiency and reproducibility of research across behavioural and brain sciences.
[44 chars over]

### 2021-11-16 application/backgroundToTheResource.tex
Updates from Overleaf

The proposed project will \textbf{enhance, reinforce and support} this key neuroinformatics
  resource to streamline \textbf{reproducible experimental design, data collection and analysis}, while making \textbf{standardised implementations of cutting-edge machine-learning methodologies widely available to the community}, potentially creating the most powerful experimental
control and analysis tool available to neuroscientists, psychologists and
ethologists worldwide.

>>>>

The proposed project will \textbf{enhance, reinforce and support} this key neuroinformatics
resource to streamline \textbf{reproducible experimental design, data collection and analysis}, while making \textbf{standardised implementations of cutting-edge machine-learning methodologies} widely available to the community, potentially creating the most powerful experimental
control and analysis tool available to neuroscientists, psychologists and
ethologists worldwide.

### 2021-11-16 application/caseForSupport.tex
11/16 changes Maneesh
Added package float and fixed problem with 11point
Updates from Overleaf

\section{Resource management}

\input resourceManagement

\section{Long term sustainability planning}

\section{Potential for economic and societal impact}
\textbf{Understanding the brain.} Systems neuroscientists aim to uncover how adaptive behaviour arises
from computation in neural circuits in different brain regions. Extending the Bonsai ecosystem will facilitate experiments that connect neural circuits with elemental behavioural processes. Long-term, this research will help identify the missing links between circuit dysfunction and behavioural deficits in neurological or mental health disorders, and will pave the way for the development of new neural circuit biomarkers and therapies. The Bonsai ecosystem will also enable the experimental testing of emerging theories of brain function and learning, and thus contribute to the neuro-inspired machine learning algorithms that fuel intelligent technologies, increasingly important in our society.

@@

>>>>

% \section{Resource management}

\input resourceManagementandSustainability

% \section{Long term sustainability planning}

\section{Potential for economic and societal impact}
\textbf{Understanding the brain and behaviour.} Systems neuroscientists aim to uncover how adaptive behaviour arises
from computation in neural circuits in different brain regions. Extending the Bonsai ecosystem will facilitate new types of experiment that causally connect neural circuits with elemental behavioural processes. Long-term, this research will help identify the missing links between circuit dysfunction and behavioural deficits in neurological or mental health disorders, and will pave the way for the development of new neural circuit biomarkers and therapies. The Bonsai ecosystem will also enable the experimental testing of emerging theories of brain function and learning, and thus contribute to the neuro-inspired machine learning algorithms that fuel intelligent technologies, increasingly important in our society.
The enhanced Bonsai platform will also support purely behavioural experiments, in which the animals' posture, body parts or location are tracked in real-time, allowing investigation of environmental impacts and captivit the impact of the envuronmental variables in behaviour but also enhace monitoring of intaractive experimental designs animals or other

@@

\input{trackRecord}

### 2021-11-16 application/detailsOfResource.tex
Updates from Overleaf

\section{Details of the proposed resource enhancements}

We intend to create software infrastructure to enable the\textbf{ integration of advanced data analysis methods} into
Bonsai (Section~\ref{sec:infrastructure}).
%
To demonstrate the use of this infrastructure, and to provide Bonsai users with
an initial set of advanced machine intelligence neural data analysis methods,
we will add to the Bonsai ecosystem machine learning functionality to control
neuroscience experiments and to analyse data generated by them
(Section~\ref{sec:functionality}).

@@

Currently Bonsai has its own package manager, and its community of users have
already extended Bonsai's functionality with several contributed packages.
%
However, these packages need to
be written in C\#, while most current neuroscience data analysis methods are
written in Python, R or Matlab. We propose to add to Bonsai capabilities to
communicate with software written in these languages
(Section~\ref{sec:softwareInfrastructure}).  These capabilities will allow a
large number of neuroscience data analysis methods to be easily integrated into
the Bonsai ecosystem and will provide Bonsai users a large repertoire of
advanced machine learning methods for their behavioural and neural data
analysis.
%
Using these capabilities we will integrate into Bonsai an initial set of
state-of-the art behavioural and neural data analysis methods
(Section~\ref{sec:functionality}).

@@

\subsubsection{Generality}
\label{sec:generality}

@@

\subsubsection{Extensibility}
\label{sec:extensibility}

To incorporate machine learning functionality into the Bonsai ecosystem, we
could
implement this functionality directly into Bonsai, as done in
\texttt{scikit-learn}. Alternatively, or in addition, we could build software
infrastructure to facilitate the integration of existing machine learning
software into Bonsai. We chose to begin with the second alternative because
high-quality implementations of machine learning functionality already exist,
which can be quickly integrated into the Bonsai ecosystem. In addition, the
second alternative is more sustainable, as it will allow the continued addition
of new machine learning functionality to Bonsai after the completion of the
BBSRC-BBR funding period. Later we may decide to implement some widely-used
machine learning functionality directly into Bonsai (e.g., for performance or
robustness issues).

Currently Bonsai has its own package manager, and its community of users have
already extended Bonsai's functionality with several contributed packages
(e.g., \ldots). However, these packages need to be written in C\#, while most
current data analysis methods are written in Python, R or Matlab. We propose to
add to Bonsai capabilities to communicate with software written in these
languages.

Fortunately, there exist software that allow C\# program
to call and be called by programs in these other languages. For the
communication of C\# programs with Python we will use
Python.NET\footnote{\href{https://github.com/pythonnet/pythonnet}{https://github.com/pythonnet/pythonnet}},
with R programs we will use
R.Net\footnote{\href{https://github.com/rdotnet/rdotnet}{https://github.com/rdotnet/rdotnet}}
and with Matlab we will use the built in Matlab .NET
interface\footnote{\href{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}}.
In cases where these software cannot support some type of communication between
C\# and another language we will use the messaging library
ZeroMQ\footnote{\href{https://zeromq.org/}{https://zeromq.org/}}.

>>>>

\section{Details of proposed resource enhancements}

The core proposed developments will enhance Bonsai's extensibility in two ways.
%
First, we will create a new package infrastructure to enable the integration of software toolboxes implemented in non-native languages into Bonsai.
%
Second, we will create a new general abstraction framework within Bonsai suited to the implementation and application of data analytic algorithms.
%
Both frameworks will expose public interfaces available to all developers within the user community, thus simplifying and accelerating community-driven development.

Furthermore, we will build on these extensibility developments ourselves to implement a suite of advanced machine intelligence neural data analysis methods that will be made available as one or more core Bonsai packages.

\subsection{Extensions to incorporate non-native code packages into Bonsai}

As it stands, Bonsai is designed to be extensible and is distributed with a package management framework that controls installation of both optional core packages and user-contributed extensions.
%
However, the present framework requires extensions to be written in Bonsai's native C\# programming language.
%
This specificity restricts the number of potential contributors to the ecosystem.
%
It also hinders easy incorporation of published analytic tools, many of which are implemented for interpreted language systems such as Python, R, or MATLAB.

We will extend the Bonsai package interface to allow direct interaction with the runtime execution environments that implement these common languages.  
%
Where possible, the extensions will be based on existing pipeline tools that allow C\#-based programs to interact with a number of alternative run-time engines.
%
These include:
%
Python.NET\footnote{\href{https://github.com/pythonnet/pythonnet}{https://github.com/pythonnet/pythonnet}},
%
R.Net\footnote{\href{https://github.com/rdotnet/rdotnet}{https://github.com/rdotnet/rdotnet}}
%
and the native .NET interface shipped with MATLAB\footnote{\href{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}{https://uk.mathworks.com/help/matlab/matlab\_external/using-net-from-matlab-an-overview.html}}.
%
If necessary, we will also incorporate direct inter-runtime communication using the messaging library ZeroMQ\footnote{\href{https://zeromq.org/}{https://zeromq.org/}}.

\subsection{General abstraction framework for data processing}
\label{sec:generality}

a package management , and its community of users have
already extended Bonsai's functionality with several contributed packages.
%
However, these packages need to
be written in C\#, while most current neuroscience data analysis methods are
written in Python, R or Matlab. We propose to add to Bonsai capabilities to
communicate with software written in these languages
(Section~\ref{sec:softwareInfrastructure}).  These capabilities will allow a
large number of neuroscience data analysis methods to be easily integrated into
the Bonsai ecosystem and will provide Bonsai users a large repertoire of
advanced machine learning methods for their behavioural and neural data
analysis.
%
Using these capabilities we will integrate into Bonsai an initial set of
state-of-the art behavioural and neural data analysis methods
(Section~\ref{sec:functionality}).

% \subsubsection{Extensibility}
% \label{sec:extensibility}

% To incorporate machine learning functionality into the Bonsai ecosystem, we
% could
% implement this functionality directly into Bonsai, as done in
% \texttt{scikit-learn}. Alternatively, or in addition, we could build software
% infrastructure to facilitate the integration of existing machine learning
% software into Bonsai. We chose to begin with the second alternative because
% high-quality implementations of machine learning functionality already exist,
% which can be quickly integrated into the Bonsai ecosystem. In addition, the
% second alternative is more sustainable, as it will allow the continued addition
% of new machine learning functionality to Bonsai after the completion of the
% BBSRC-BBR funding period. Later we may decide to implement some widely-used
% machine learning functionality directly into Bonsai (e.g., for performance or
% robustness issues).

@@

@@

@@

### 2021-11-16 application/trackRecord.tex
Updates from Overleaf

The team responsible for the proposed project comprises two academic
institutes, the Gatsby Computational Neuroscience Unit (GCNU) and the Sainsbury
Wellcome Centre for Neural Circuits and Behavior (SWC), and a private
collaborator, NeuroGEARS Ltd.

>>>>

The team responsible for the proposed project comprises two academic
institutes at University College London, the Gatsby Computational Neuroscience Unit (GCNU) and the Sainsbury
Wellcome Centre for Neural Circuits and Behavior (SWC), and a private
collaborator, NeuroGEARS Ltd.
